---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file, then run devtools::build_readme(). -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# baselinr

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/baselinr)](https://CRAN.R-project.org/package=baselinr)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/baselinr)](https://CRAN.R-project.org/package=baselinr)
[![R-CMD-check](https://github.com/zl1212-ship-it/baselinr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/zl1212-ship-it/baselinr/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![DOI](https://zenodo.org/badge/1279288173.svg)](https://doi.org/10.5281/zenodo.20937887)
<!-- badges: end -->

`baselinr` builds report-ready **baseline equivalence** tables for impact
evaluations in education research, following the conventions of the
[What Works Clearinghouse (WWC)](https://ies.ed.gov/ncee/wwc/). Given a treatment
indicator and a set of covariates, it reports the appropriate standardized effect
size for each covariate, **Hedges' g** for continuous covariates and the
**Cox index** for binary ones, together with the WWC equivalence category.

It is a thin, education-specific reporting layer. For general-purpose covariate
balance assessment, see [`cobalt`](https://ngreifer.github.io/cobalt/);
`baselinr` focuses narrowly on the WWC equivalence categories that education
evaluation reports are required to state.

## Installation

Install the released version from CRAN:

```r
install.packages("baselinr")
```

Or the development version from GitHub:

```r
# install.packages("remotes")
remotes::install_github("zl1212-ship-it/baselinr")
```

## Example

```{r}
library(baselinr)

study <- data.frame(
  treat   = c(1, 1, 1, 0, 0, 0),
  pretest = c(5, 6, 7, 4, 5, 6), # continuous -> Hedges' g
  female  = c(1, 0, 1, 0, 0, 1)  # binary     -> Cox index
)

knitr::kable(baseline_equivalence(study, treatment = "treat"), digits = 3)
```

The WWC categories are:

| Effect size (absolute) | Category | Meaning |
|---|---|---|
| `<= 0.05` | `satisfied` | Baseline equivalence holds. |
| `0.05`–`0.25` | `satisfied_with_adjustment` | Holds only if the covariate is adjusted for in the impact model. |
| `> 0.25` | `not_satisfied` | Cannot establish equivalence. |

## Visualise and format

`love_plot()` shows the standardized effect size of every covariate against the
WWC thresholds (requires `ggplot2`):

```{r loveplot, fig.alt = "Love plot of standardized effect sizes by covariate", fig.width = 7, fig.height = 2.6, dpi = 150}
love_plot(baseline_equivalence(study, treatment = "treat"))
```

`gt_baseline()` renders the same table as a formatted `gt` table for reports and
Quarto/HTML (requires `gt`):

```r
gt_baseline(baseline_equivalence(study, treatment = "treat"))
```

## Scope

Continuous covariates use Hedges' g (with the WWC small-sample correction);
binary covariates (numeric `0/1`, logical, or two-level factor) use the WWC Cox
index. `wwc_summary()` collapses the table into an overall verdict, and
`attrition()` reports overall and differential attrition. See `NEWS.md` for the
roadmap.

## License

MIT © Yuxia Liang
